---
title: "Statistical Analysis Plan for title"
subtitle: 'DOCUMENT: SAP-2021-004-TV-v01'
author: '**From:** Felipe Figueiredo **To:** Thiago Vivacqua'
date: '**Date:** yyyy-mm-dd'
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
library(philsfmisc)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', Inf)
```

---

**Document version**

```{r, echo = FALSE}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

<!-- # Assinaturas -->

<!-- ```{r, echo=FALSE} -->
<!-- sig.field <- "__________________________" -->
<!-- date.field <- "_____________" -->
<!-- Stat <- c("Elaborador", "Nome", "Função", sig.field, date.field) -->
<!-- Reviewer <- c("Revisado por", "", "", sig.field, date.field) -->
<!-- Approver <- c("Verificado por", "", "", sig.field, date.field) -->
<!-- Final.Approver <- c("Aprovação final", "", "", sig.field, date.field) -->

<!-- sigs <- rbind( -->
<!--   Stat -->
<!--   , Reviewer -->
<!--   , Approver -->
<!--   , Final.Approver -->
<!--   ) -->
<!-- rownames(sigs) <- NULL -->
<!-- colnames(sigs) <- c("Papel", "Nome", "Função", "Assinatura", "Data") -->

<!-- # pander(sigs, split.cells = c(9, 14, 14, 16, 8), split.table = Inf) -->
<!-- kable(sigs) -->
<!-- ``` -->

# Abbreviations

- BMI: body mass index
- LCL: lateral collateral ligament
- LET: lateral extra-articular tenodesis
- MCL: medial collateral ligament
- PCL: posterior cruciate ligament

# Introduction

## Context

## Objectives

## Hypotheses

# Data cleaning

## Raw data

The raw data was provided in an Excel file with two sheets, one for each group (LET and No LET).
As a first step of pre-processing the raw data was copied into a new sheet with both groups, with one observation (patient) per line.
The column structure of the resulting "clean" table corresponds to the original per-group tables.

Variables in the raw datasheet were grouped into sections.
For this analysis only the baseline **demographic** characteristics and the variables in the **surgical data most recent revision ACLR** section were selected.

Column names were adjusted to avoid repeated names, using suffixes to identify the section each variable originally came from, as follows:

- section `concomitant injuries primary ACLR` - received the `pri.inj` suffix
- section `Concomitant injuries most recent revision ACLR` - received the `inj` suffix
- section `surgical data primary ACLR` - received the `pri` suffix
- section `surgical data most recent revision ACLR` - names were kept as the original

Some dates in `Date of last clinical visit` column (`Follow-up notes` data section) were typed with different separator symbols, and all values in this column were standardized to allow date format recognition.

There were two columns named `grade` (`Slope` data section), and they were both renamed `grade1` and `grade2`.

The original raw data provided descriptive information for the encoding of each variable in the first line.
A Data Dictionary was created from this information and was saved in a separate table for ease of consultation and processing.

The resulting "clean" raw data sheet was shared with the client for archival purposes.

## Analytic dataset

After the cleaning process the following variables were included in the analysis:

- id
- age
- sex
- bmi
- let
- stage
- graft
- graft_diameter
- femoral_fixation
- tibial_fixation
- medial_meniscus
- lateral_meniscus
- cartilage
- mcl
- lcl
- pcl
- implant_hto
- femoral_tunnel
- tibial_tunnel

All variables in the analytic set were labelled according to the raw data provided and values were labelled according to the data dictionary for the preparation of production-quality results tables and figures.

The analytic dataset will be included in the private version of the report, and will be omitted from the public version of the report.

# Study variables

## Primary and secondary outcomes

The primary outcome of this analysis is the difference in surgery failure between the two study groups.
The secondary outcomes are the differences between the baseline demographic variables Age, Sex and BMI.

## Covariates

# Statistical methods

## Statistical analyses

### Descriptive analyses

### Inferential analyses

### Statistical modeling

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be two-tailed.

## Study size and Power

### Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Exceptions and Observations

# References

# Appendix

## Availability

This document can be downloaded in the following address:

`https://github.com/philsf-biostat/SAR-2021-004-TV/`
