---
title: "Analytical Plan for Comparison of baseline characteristics of patients undergoing ACLR under usage of the LET surgical technique"
subtitle: 'DOCUMENT: SAP-2021-004-TV-v01'
author: '**From:** Felipe Figueiredo **To:** Thiago Vivacqua'
date: '**Date:** 2021-08-24'
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
library(philsfmisc)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', Inf)
```

---

**Document version**

```{r, echo = FALSE}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

# Abbreviations

- ACL: anterior cruciate ligament
- ACLR: ACL revision
- BMI: body mass index
- LCL: lateral collateral ligament
- LET: lateral extra-articular tenodesis
- MCL: medial collateral ligament
- MTPS: medial tibial posterior slope
- PCL: posterior cruciate ligament

# Introduction

## Context

## Objectives

Compare the baseline characteristics of patients undergoing ACL revision (ACLR) surgery whether or not they the LET technique was employed in the ACLR.

## Hypotheses

LET and no LET groups are comparable at baseline.

# Data cleaning

```{r cleaning, include=FALSE}
source("scripts/input.R", encoding = 'UTF-8')
```

## Raw data

The raw data was provided in an Excel file with two sheets, one for each group (LET and No LET).
As a firsts processing step the raw data was copied into a new sheet with both groups, with one observation (patient) per line.
The column structure of the resulting "clean" table corresponds to the original per-group tables.

In total there were 126 variables in the dataset received.
Variables in the raw data sheet were grouped into sections.
For this analysis only the baseline **demographic** characteristics and the variables in the **surgical data most recent revision ACLR** section were selected.

Column names were adjusted to avoid repeated names, using suffixes to identify the section each variable originally came from, as follows:

- section `concomitant injuries primary ACLR` -- received the `pri.inj` suffix
- section `Concomitant injuries most recent revision ACLR` -- received the `inj` suffix
- section `surgical data primary ACLR` -- received the `pri` suffix
- section `surgical data most recent revision ACLR` -- names were kept as the original

Some dates in `Date of last clinical visit` column (`Follow-up notes` data section) were typed with different separator symbols, and all values in this column were standardized to allow date format recognition.

There were two columns named `grade` (`Slope` data section), and they were both renamed `grade1` and `grade2`.

The original raw data provided descriptive information for the encoding of each variable in the first line.
A Data Dictionary was created from this information and was saved in a separate table for ease of consultation and processing.

The resulting "clean" raw data sheet was shared with the client for archival purposes.

## Analytical dataset

All variables in the analytical set were labeled according to the raw data provided and values were labeled according to the data dictionary for the preparation of production-quality results tables and figures.

After the cleaning process `r ncol(analytical)` variables were included in the analysis with `r nrow(analytical)` observations.
Table 1 shows the structure of the analytical dataset.

```{r}
analytical_mockup %>%
  pander(caption = "**Table 1** Analytical dataset structure after variable selection and cleaning.")
```


The analytical dataset will be included in the private version of the report, and will be omitted from the public version of the report.

# Study variables

## Primary and secondary outcomes

This analysis describes the epidemiological characteristics of surgical patients that undergone ACLR, and no endpoints will be assessed.

## Covariates

The distribution of the characteristics will be compared at baseline between the two study groups.
Variables to be considered for baseline comparison are:

- Age
- Sex
- BMI
- Graft diameter
- Medial meniscus
- Lateral meniscus
- Cartilage

# Statistical methods

## Statistical analyses

### Descriptive analyses

The epidemiological profile of each study group (LET and no LET) will be described at baseline.
Demographic and clinical variables will be described as mean and standard deviation or as counts and proportions, as appropriate.
Distributions will be summarized in tables and visualized in exploratory plots.

### Inferential analyses

All comparisons will be performed as univariate analyses.
Continuous variables will be compared between groups with the independent t test with Welch correction.
Difference in distribution of categorical variables will be assessed with the Fisher exact test.

### Statistical modeling

No adjustment for covariates will be performed.

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be two-tailed.

## Study size and Power

Raw data received has 70 observations corresponding to patients assessed at primary ACL surgery and ACL revision.
Surgeries that employed the LET technique include 38 patients and 32 patients undergone surgery with LET technique.

With these group sizes a t test can detect an effect size as large as
**d = `r format.float(pwr::pwr.t2n.test(n1 = 38, n2 = 32, power = .8, sig.level = .05)$d)`**
with 80% power and 5% significance level (Cohen, 1988).

### Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Exceptions and Observations

N/A.

# References

- **SAR-2021-004-TV-v01** -- Comparison of baseline characteristics of patients undergoing ACLR under usage of the LET surgical technique.
- Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge.

# Appendix

## Availability

Both this analytical plan and the corresponding analysis report (**SAR-2021-004-TV-v01**) can be downloaded in the following address:

`https://github.com/philsf-biostat/SAR-2021-004-TV/`
