---
title: "Epidemiological profile of patients undergoing ACLR and comparison of baseline characteristics between usage of the LET surgical technique"
subtitle: 'DOCUMENT: SAR-2021-004-TV-v01'
author: '**From:** Felipe Figueiredo **To:** Thiago Vivacqua'
date: '**Date:** yyyy-mm-dd'
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAR_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
library(philsfmisc)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', Inf)
```

---

**Document version**

```{r, echo = FALSE}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

<!-- # Assinaturas -->

<!-- ```{r, echo=FALSE} -->
<!-- sig.field <- "__________________________" -->
<!-- date.field <- "_____________" -->
<!-- Stat <- c("Elaborador", "Nome", "Função", sig.field, date.field) -->
<!-- Reviewer <- c("Revisado por", "", "", sig.field, date.field) -->
<!-- Approver <- c("Verificado por", "", "", sig.field, date.field) -->
<!-- Final.Approver <- c("Aprovação final", "", "", sig.field, date.field) -->

<!-- sigs <- rbind( -->
<!--   Stat -->
<!--   , Reviewer -->
<!--   , Approver -->
<!--   , Final.Approver -->
<!--   ) -->
<!-- rownames(sigs) <- NULL -->
<!-- colnames(sigs) <- c("Papel", "Nome", "Função", "Assinatura", "Data") -->

<!-- # pander(sigs, split.cells = c(9, 14, 14, 16, 8), split.table = Inf) -->
<!-- kable(sigs) -->
<!-- ``` -->

```{r results, include=FALSE}
source("scripts/results.R", encoding = 'UTF-8')
```

# Abbreviations

- ACL: anterior cruciate ligament
- ACLR: ACL revision
- BMI: body mass index
- LCL: lateral collateral ligament
- LET: lateral extra-articular tenodesis
- MCL: medial collateral ligament
- PCL: posterior cruciate ligament

# Introduction

## Context

## Objectives

Compare the baseline characteristics of patients undergoing ACL revision (ACLR) surgery whether or not they the LET technique was employed in the ACLR.

## Hypotheses

LET and no LET groups are comparable at baseline.

## Data reception and cleaning

In total there were 126 variables in the dataset received.
Variables in the raw data sheet were grouped into sections.
For this analysis only the baseline **demographic** characteristics and the variables in the **surgical data most recent revision ACLR** section were selected.

All variables in the analytical set were labeled according to the raw data provided and values were labeled according to the data dictionary for the preparation of production-quality results tables and figures.

After the cleaning process `r ncol(analytical)` variables were included in the analysis with `r nrow(ncol)` observations.
Table 1 shows the structure of the analytical dataset.

# Methods

## Study Variables

### Primary and secondary outcomes

This analysis describes the epidemiological characteristics of surgical patients that undergone ACLR, and no endpoints will be assessed.

### Covariates

The distribution of the characteristics will be compared at baseline between the two study groups.
Variables to be considered for baseline comparison are:

- Age
- Sex
- BMI
- Graft diameter
- Medial meniscus
- Lateral meniscus
- Cartilage

## Statistical analyses

The epidemiological profile of each study group (LET and no LET) will be described at baseline.
Demographic and clinical variables will be described as mean and standard deviation or as counts and proportions, as appropriate.
Distributions will be summarized in tables and visualized in exploratory plots.

All comparisons will be performed as univariate analyses.
Continuous variables will be compared between groups with the independent t test with Welch correction.
Difference in distribution of categorical variables will be assessed with the Fisher exact test.
No adjustment for covariates will be performed.

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be two-tailed.

### Statistical packages

This analysis was performed using statistical software `R` version `r getRversion()`.

# Results

## Study population and follow up

## Inferential analysis

**Table 1** Patient characteristics at baseline.
1 n (%); Mean (SD) 
p-value: Fisher's exact test; Welch Two Sample t-test 

```{r inf_baseline}
inf_baseline %>% as_kable(align = "lcccccc")
```



# Exceptions and Observations

# Conclusions

# References

- **SAP-2021-004-TV-v01** -- Statistical Analysis Plan for Epidemiological profile of patients undergoing ACLR and comparison of baseline characteristics between usage of the LET surgical technique
- Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge.

# Appendix

## Exploratory data analysis

## Availability

This document can be downloaded in the following address:

`https://github.com/philsf-biostat/SAR-2021-004-TV/`

## Analytical dataset

Due to confidentiality the data-set used in this analysis cannot be shared online in the public version of this report.

```{r data}
# print analytical of analytical_mockup whether it is the private or public version of the SAR
# analytical %>%
analytical_mockup %>%
  pander(caption = "**Table A1** Analytical dataset")
```
