---
title: "title"
subtitle: 'DOCUMENT: SAR-2021-004-TV-v01'
author: '**From:** Felipe Figueiredo **To:** Thiago Vivacqua'
date: '**Date:** yyyy-mm-dd'
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAR_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
library(philsfmisc)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', Inf)
```

---

**Document version**

```{r, echo = FALSE}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

---

<!-- # Assinaturas -->

<!-- ```{r, echo=FALSE} -->
<!-- sig.field <- "__________________________" -->
<!-- date.field <- "_____________" -->
<!-- Stat <- c("Elaborador", "Nome", "Função", sig.field, date.field) -->
<!-- Reviewer <- c("Revisado por", "", "", sig.field, date.field) -->
<!-- Approver <- c("Verificado por", "", "", sig.field, date.field) -->
<!-- Final.Approver <- c("Aprovação final", "", "", sig.field, date.field) -->

<!-- sigs <- rbind( -->
<!--   Stat -->
<!--   , Reviewer -->
<!--   , Approver -->
<!--   , Final.Approver -->
<!--   ) -->
<!-- rownames(sigs) <- NULL -->
<!-- colnames(sigs) <- c("Papel", "Nome", "Função", "Assinatura", "Data") -->

<!-- # pander(sigs, split.cells = c(9, 14, 14, 16, 8), split.table = Inf) -->
<!-- kable(sigs) -->
<!-- ``` -->

# Abbreviations

# Introduction

## Objectives

## Data reception and cleaning

# Methods

```{r results, include=FALSE}
source("scripts/results.R", encoding = 'UTF-8')
```

## Variables

### Primary and secondary outcomes

### Covariates

## Statistical analyses

### Statistical packages

This analysis was performed using statistical software `R` version `r getRversion()`.

# Results

## Study population and follow up

## Inferential analysis

# Exceptions and Observations

# Conclusions

# References

# Appendix

## Exploratory data analysis

## Availability

This document can be downloaded in the following address:

`https://github.com/philsf-biostat/SAR-2021-004-TV/`

## Analytical data-set

Due to confidentiality the data-set used in this analysis cannot be shared online in the public version of this report.

```{r data}
# pander(analytic)
```
